\sysfull

### 2017-11-17 paper/abstract.tex

Computation consistency on intermittently-powered devices can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections. Results so far suggest that a task-based approach performs better than checkpointing. However, existing tasks-based approaches do not adapt to changing energy levels and storage. This makes the code inefficient (when executed on a larger energy buffer than intended) or can potentially completely disable the program (in the opposite case). Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing. If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve execution times. \todo{Provide concrete numbers for improvement}{Amjad} \todo{Provide concrete numbers for checkpointing improvement}{Amjad}

>>>>

Computation consistency on intermittently-powered devices can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections. Results so far suggest that a task-based approach performs better than checkpointing. However, existing tasks-based approaches do not adapt to changing energy levels and storage. This makes the code inefficient (when executed on a larger energy buffer than intended) or can potentially completely disable the program (in the opposite case). Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing. If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve execution times up to 4.5 times compared to static task code.

### 2017-11-17 paper/background.tex
background: add a paragraph break
rename the fram access figures

\begin{figure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
	\end{subfigure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
		\caption{Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
    \end{subfigure}
	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}\label{fig:code_demo_incosistency}
\end{figure}

@@

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}                             & Reg. + Stack     \\
		DINO~\cite{dino}                                     & Reg. + WAR variables \\
		Chain~\cite{chain}                                   & PC   + Channel data\\
		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
		\hline
	\end{tabular}
	\caption{Memory consistency enforcement overheads of intermittent execution models. \emph{Reg.} includes the entire register file, \emph{PC} is the program counter, \emph{Channel data} are variables explicitly task-shared by the programmer, and \emph{WAR variables} are variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

@@

Task-based models incur costs that motivate \sys.

@@

\begin{figure}
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_write.eps}
	\caption{Cost of \emph{write} operation.}
\end{subfigure}%
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_read.eps}
	\caption{Cost of \emph{read} operation.}
\end{subfigure}
	\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write and read operation.\todo{pico jole$\rightarrow$Picojoule}{Amjad}}\label{fig:framEnergy}
\end{figure}

@@

\textbf{Fixed-size Tasks are Inflexible and Inefficient.} If a programmer writes a task that consumes {\em more} energy than the device can buffer, the task will never complete using buffered energy. {\em preventing forward progress} by repeatedly re-executing. If a programmer writes tasks that consume far {\em less} energy than a device can buffer, the system may operate {\em inefficiently}. When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit). However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity. Buffer sizes from prior work vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer. The key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer. \sys's {\em task coalescing}, described
in Section~\ref{sec:task_coalescing}, is motivated by these observations.

>>>>

\begin{figure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
	\end{subfigure}
	\begin{subfigure}[t]{\linewidth}
		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
    \end{subfigure}
	\caption{\textbf{Code example demonstrating effect of write after read on volatile memory checkpointing.}}\label{fig:code_demo_incosistency}
\end{figure}

@@

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}                             & Reg. + Stack     \\
		DINO~\cite{dino}                                     & Reg. + WAR variables \\
		Chain~\cite{chain}                                   & PC   + Channel data\\
		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
		\hline
	\end{tabular}
	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

@@

%Task-based models incur costs that motivate \sys.

@@

\begin{figure}
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
	\caption{Cost of \emph{write} operation}
\end{subfigure}%
\begin{subfigure}[t]{.49\columnwidth}
	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
	\caption{Cost of \emph{read} operation}
\end{subfigure}
	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.}}\label{fig:framEnergy}
\end{figure}

@@

\textbf{Fixed-size Tasks are Inflexible and Inefficient.} If a programmer writes a task that consumes {\em more} energy than the device can buffer, the task will never complete using buffered energy {\em preventing forward progress} by repeatedly re-executing. If a programmer writes tasks that consume far {\em less} energy than a device can buffer, the system may operate {\em inefficiently}. When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit). However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions.

Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity. Buffer sizes from prior work vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer. The key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer. \sys's {\em task coalescing}, described
in Section~\ref{sec:task_coalescing}, is motivated by these observations.

### 2017-11-17 paper/evaluation.tex

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalStrategies}
	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'.}
	\label{fig:coalescing}
\end{figure}

>>>>

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalStrategies}
	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'. \todo{Change WHD to TSHC and HD to HC}{Amjad}}
	\label{fig:coalescing}
\end{figure}
